### 2019-09-14 ic-journal/contraction.tex
text

Milner's ``unique solution of equations'' theorem for $\wb$
(Theorem~\ref{t:Mil89})
\hl{brings a new proof technique for proving (weak) bisimilarities. However, it has
severe limitations: the equations must be guarded and sequential (plus
all sums where equation variables appear must be guarded sums).
This greatly limits the usefulness of the technique, since
the occurrences of other operators above the variables, such as parallel
composition and restriction, in general cannot be removed.}
The constraints of Theorem~\ref{t:Mil89}, however, can be greatly
weakened if we move from equations to a special kind of inequations called
  \emph{contractions}.

>>>>

Milner's ``unique solution of equations'' theorem for $\wb$
(Theorem~\ref{t:Mil89})
\hlD{brings a new proof technique for proving (weak) bisimilarities. However, it has
 limitations: the equations must be guarded and sequential (moreover
all sums where equation variables appear must be guarded sums).
This limits the usefulness of the technique, since
the occurrences of other operators above the variables, such as parallel
composition and restriction, in general cannot be removed.}
The constraints of Theorem~\ref{t:Mil89}, however, can be 
weakened if we move from equations to a special kind of inequations called
  \emph{contractions}.

### 2019-09-14 ic-journal/equa.tex
text

\hl{In the CCS syntax, 
a recursion $\recu A  P$ acts as a binder for $A$ in the body $P$. 
This gives rise, in the expected manner, to the notions of 
\emph{free} and \emph{bound} recursion variables in a CCS expression.

 given in previous section, there is the syntactic
possibility that an agent variable $A$ occurs
outside the recursion operator of the same variable (or, equivalently, a
process constant $C$ is used but undefined). We say such a variable is
\emph{free}, while those variables wrapped by their recursions are
\emph{bound}.}
\hl{A CCS \emph{tem} (or \emph{expression}, i.e. those with valid syntax)
without any free variable is called a CCS \emph{process}.}
\hl{For instance, the variable $X$ is free in $a.X + \recu Y (b.Y)$, while $Y$
is bound. There is also the syntactic possibility that the same
variable $X$ is
both free and bound, e.g. in $a.X + \recu X (b.X)$.
This setting does not cause any ambiguity, but sometimes leads to more
complex proofs.} (c.f. Section~\ref{sec:multivariate} for more details.)

\hl{In this paper (and the formalisation work), we reuse the free agent
variables as \emph{equation variables}. This eliminates the need of
another type for CCS equations, and we can also reuse the existing
variable substitution operation} (c.f. SOS rule for the Recursion in
Fig.~\ref{f:LTSCCS}) \hl{for the substitution of equation variables.
For example, the result of substituting equation variable $X$ to $\nil$ in $a.X +
\recu X (b.X)$, denoted by $(a.X + \recu X (b.X)) \sub {\nil} X$, is
$a.\nil + \recu X (b.X)$ with the part $\recu X (b.X)$
untouched. Multivariate substitutions are written in the same syntax,
e.g. $E \sub {\til P} {\til X}$. Whenever $\til X$ is clear from the
context, we may also write $E[\til P]$ instead of $E \sub {\til P} {\til
  X}$. (and $E[P]$ for $E \sub P X$ if there is a single equation variable $X$.)}

\hl{In fact, free agent variables have the same transitional behavior
as the deadlock $\nil$ according to the SOS rules, as there is
no rule for their transitions at all. In fact, most CCS theorems still hold if
the involved CCS terms contain free variables. (The most notable
exceptions are all versions of unique solution of equations/contracts,
where all solutions must be \emph{pure} processes (i.e. no free variable).)}

\subsection{Systems of equations}
\label{ss:SysEq}

Milner's ``unique solution of equations'' theorems~\cite{Mil89} intuitively
say that, if a context $C$ \hl{(i.e., a CCS expression with possibly
free variables)}\footnote{\hl{Rigorously speaking, under our setting
(i.e.~reusing free agent variables as equation variables) not all
valid CCS expressions
are valid contexts: those where free variables occur inside
recusion operators must be all excluded. For instace, $\recu X (a.X +
b.Y)$ is not a valid context with the variable $Y$. This special requirement
is perfectly aligned with CCS literature using process constants, as any
equation variable cannot occur inside the definition of any
constant. In fact, all versions of ``unique solution of
equations/contractions'' theorems do not hold if equation variables
are allowed to occur inside any recursion operator.}} obeys certain conditions,
then all processes $P$ that satisfy the equation $P \wb \ct P$ are
bisimilar with each other.

\begin{definition}[equations] % Def 3.1
  \label{def:equation}
Assume that, for each $i$ of 
 a countable indexing set $I$, we have variables $X_i$, and expressions
$E_i$ containing variables up to $\til X$. Then 
$\{ X_i = E_i\}_{i\in I}$ is 
  a \emph{system of equations}. (There is one equation $E_i$ for each variable $X_i$.)
\end{definition}

@@

\begin{definition}[solutions and unique solutions]
  \label{def:solution}
Suppose $\{ X_i = E_i\}_{i\in I}$ is a system of equations: 
\begin{itemize}
\item
 $\til P$ is a \emph{solution of the system of equations (for $\wb$)} 
if for each $i$ it holds that $P_i \wb E_i [\til P]$;
\item The system has \emph{a unique solution for $\wb$}  if whenever 
 $\til P$ and $\til Q$ are both solutions \hl{it holds that} $\til P \wb \til Q$. 
\end{itemize} 
 \end{definition}
\hl{Similarily, the \emph{(unique) solution of a system of equations for $\sim$}
(or for $\rapprox$) can be obtained by replacing all occurrences of $\wb$
in above definition with $\sim$ and $rapprox$, respectively.}

@@

\item \emph{(strongly) guarded} if, in each $E_i$, each occurrence of
  each $X_i$ is underneath a \emph{visible} prefix;

@@

\hl{In other words, if a system of equations is sequential, then for
each context $E_i$, any subcontext of $E_i$ in which $X_j $
appears, apart from $X_j$ itself, is a sum of prefixed expressions.
For instance,}
\begin{itemize}
\item $X = \tau. X + \mu . \nil$ is sequential but not 
 guarded, because the guarding prefix for the variable
is not visible;
\item $X =  \ell . X | P$ is guarded but not sequential;
\item $X =  \ell . X + \tau. \res a (a .\outC b | a.\nil)$, as well as
$X = \tau . (a. X + \tau . b .X + \tau  )$ are both guarded and sequential.
\end{itemize}

\hl{Milner has totally three versions of ``unique solution of equations''
theorems, for $\sim$, $\wb$ and $\rapprox$, respectively, but only the
following two versions are explicitly mentioned} in~\citep[p.~103, 158]{Mil89}:
\begin{theorem}[unique solution of equations for $\sim$]
\label{t:Mil89s1}
Let $E_i$ be weakly guarded with free variables up to $\til X$,
and let ${\til P} \sim {\til E}\{\til P /\til X\}$,
  ${\til Q} \sim {\til E}\{\til Q /\til X\}$. Then ${\til P} \sim {\til Q}$.
\end{theorem}

@@

\hl{The version of Milner's unique-solution theorem for $\wb$ further requires
that all sums where equation variables appear are guarded
sums}\footnote{But if the CCS syntax were defined with only guarded
  sums, i.e., $\sum_{i\in I} \mu_i.P_i$ as
  in~\cite{sangiorgi2015equations}, this addition
  requirement disappears automatically, and we can even say $\wb$ is indeed a
  congruence.}:
\begin{theorem}[unique solution of equations for $\wb$]
\label{t:Mil89}
Let $E_i$ be guarded (with guarded sums only) and sequential with free
variables up to $\til X$, and let ${\til P} \wb {\til E}\{\til P /\til X\}$,
  ${\til Q} \wb {\til E}\{\til Q /\til X\}$. Then ${\til P} \wb {\til Q}$.
\end{theorem}

The proof of \hl{the last theorem above} exploits an invariance
property on immediate \hl{derivations}
of guarded and sequential expressions, and then extracts a bisimulation
(up to bisimilarity) out
of the solutions of the system.
To see the need of the sequentiality  condition, consider
 the equation (from \cite{Mil89}) $X = \res a (a. X | \outC a)$
where $X$ is guarded but not sequential. Any process that does not use
$a$ is a solution\hl{, e.g. $\nil$ and $b.\nil$.}

\hl{For more details of above three theorems,} c.f. Section~\ref{ss:part2}
for their formalised versions (the univariate case) and
Section~\ref{sec:multivariate} for the multivariate case
(Theorem~\ref{t:Mil89s1} only).

>>>>

\hlD{In the CCS syntax, 
a recursion $\recu A  P$ acts as a binder for $A$ in the body $P$. 
This gives rise, in the expected manner, to the notions of 
\emph{free} and \emph{bound} recursion variables in a CCS expression. 
For instance,  $X$ is free in $a.X + \recu Y (b.Y)$ while $Y$
is bound; 
in $a.X + \recu X (b.X)$,
 $X$ is
both free and bound.
%  given in previous section, there is the syntactic
% possibility that an agent variable $A$ occurs
% outside the recursion operator of the same variable (or, equivalently, a
% process constant $C$ is used but undefined). We say such a variable is
% \emph{free}, while those variables wrapped by their recursions are
% \emph{bound}.}
A tem
without free variables  is a  \emph{process}.}
% This setting does not cause any ambiguity, but sometimes leads to more
% complex proofs.} (c.f. Section~\ref{sec:multivariate} for more details.)

\hlD{In this paper (and the formalisation work), we use the agent
variables also as \emph{equation variables}. This eliminates the need of
another type for  equations, and we can reuse the existing
variable substitution operation} (c.f. SOS rule for the Recursion in
Fig.~\ref{f:LTSCCS}) \hlD{for the substitution of equation variables.
For example, the result of substituting variable $X$ with $\nil$ in $a.X +
\recu X (b.X)$,  written $(a.X + \recu X (b.X)) \sub {\nil} X$, is
$a.\nil + \recu X (b.X)$ (with the part $\recu X (b.X)$
untouched). Multivariate substitutions are written in the same syntax,
e.g. $E \sub {\til P} {\til X}$. Whenever $\til X$ is clear from the
context, we may also write $E[\til P]$ instead of $E \sub {\til P} {\til
  X}$ (and $E[P]$ for $E \sub P X$ if there is a single equation variable $X$.)}

% \hl{In fact, free agent variables have the same transitional behavior
% as the deadlock $\nil$ according to the SOS rules, as there is
% no rule for their transitions at all. In fact, most CCS theorems still hold if
% the involved CCS terms contain free variables. (The most notable
% exceptions are all versions of unique solution of equations/contracts,
% where all solutions must be \emph{pure} processes (i.e. no free variable).)}

\subsection{Systems of equations}
\label{ss:SysEq}
\hlD{When discussing equations it is standard to talk about `context'. This is a 
 a CCS expression  possibly containing  free variables that, however, may not occur within
the body of recursive definitions. 
Milner's ``unique solution of equations'' }  theorems~\cite{Mil89} intuitively
\hlD{say that, if a context $C$
%  \hl{(i.e., a CCS expression with possibly
% free variables)}\footnote{\hl{Rigorously speaking, under our setting
% (i.e.~reusing free agent variables as equation variables) not all
% valid CCS expressions
% are valid contexts: those where free variables occur inside
% recusion operators must be all excluded. For instace, $\recu X (a.X +
% b.Y)$ is not a valid context with the variable $Y$. This special requirement
% is perfectly aligned with CCS literature using process constants, as any
% equation variable cannot occur inside the definition of any
% constant. In fact, all versions of ``unique solution of
% equations/contractions'' theorems do not hold if equation variables
% are allowed to occur inside any recursion operator.}} 
obeys certain conditions,
then all processes $P$ that satisfy the equation $P \wb \ct P$ are
bisimilar with each other.
}

\begin{definition}[equations] % Def 3.1
  \label{def:equation}
Assume that, for each $i$ of 
 a countable indexing set $I$, we have variables $X_i$, and expressions
$E_i$ \hlD{possibly containing such variables $\cup_i \{ X_i\}$}. Then 
$\{ X_i = E_i\}_{i\in I}$ is 
  a \emph{system of equations}. (There is one equation $E_i$ for each variable $X_i$.)
\end{definition}

@@

\begin{definition}[solutions and unique solutions]
  \label{def:solution}
Suppose $\{ X_i = E_i\}_{i\in I}$ is a system of equations: 
\begin{itemize}
\item
 $\til P$ is a \emph{solution of the system of equations (for $\wb$)} 
if for each $i$ it holds that $P_i \wb E_i [\til P]$;
\item The system has \emph{a unique solution for $\wb$}  if whenever 
 $\til P$ and $\til Q$ are both solutions \hlD{then} $\til P \wb \til Q$. 
\end{itemize} 
 \end{definition}
\hl{Similarily, the \emph{(unique) solution of a system of equations for $\sim$}
(or for $\rapprox$) can be obtained by replacing all occurrences of $\wb$
in above definition with $\sim$ and $\rapprox$, respectively.}

@@

\item \hlD{\emph{guarded}} if, in each $E_i$, each occurrence of
  each $X_i$ is underneath a \emph{visible} prefix;

\finish{i have removed 
``(strongly)''}

@@

\hlD{In other words, if a system of equations is sequential, then for
each  $E_i$, any subexpression of $E_i$ in which $X_j $
appears, apart from $X_j$ itself, is a sum of prefixed expressions.
For instance,}
\begin{itemize}
\item $X = \tau. X + \mu . \nil$ is sequential but not 
 guarded, because the guarding prefix for the variable
is not visible;
\item $X =  \ell . X | P$ is guarded but not sequential;
\item $X =  \ell . X + \tau. \res a (a .\outC b | a.\nil)$, as well as
$X = \tau . (a. X + \tau . b .X + \tau  )$ are both guarded and sequential.
\end{itemize}

\hlD{Milner has  three versions of  ``unique solution of equations''
theorems, for $\sim$, $\wb$ and $\rapprox$, respectively, though only the
following two versions are explicitly mentioned} in~\citep[p.~103, 158]{Mil89}:
\begin{theorem}[unique solution of equations for $\sim$]
\label{t:Mil89s1}
Let $E_i$ be weakly guarded with free variables in $\til X$,
and let ${\til P} \sim {\til E}\{\til P /\til X\}$,
  ${\til Q} \sim {\til E}\{\til Q /\til X\}$. Then ${\til P} \sim {\til Q}$.
\end{theorem}

@@

\hlD{The version of Milner's unique-solution theorem for $\wb$ further requires
that all sums are guarded}:
% \footnote{But if the CCS syntax were defined with only guarded
%   sums, i.e., $\sum_{i\in I} \mu_i.P_i$ as
%   in~\cite{sangiorgi2015equations}, this addition
%   requirement disappears automatically, and we can even say $\wb$ is indeed a
%   congruence.}:
\begin{theorem}[unique solution of equations for $\wb$]
\label{t:Mil89}
Let $E_i$ be \hlD{guarded}  and sequential with free
variables up to $\til X$, and let ${\til P} \wb {\til E}\{\til P /\til X\}$,
  ${\til Q} \wb {\til E}\{\til Q /\til X\}$. Then ${\til P} \wb {\til Q}$.
\end{theorem}

\finish{below and elsewhere: the correct terminology is ``derivative'', instead of ``derivation''}

The proof of \hlD{the  theorem above} exploits an invariance
property on immediate \hlD{derivatives}
of guarded and sequential expressions, and then extracts a bisimulation
(up to bisimilarity) out
of the solutions of the system.
To see the need of the sequentiality  condition, consider
 the equation (from \cite{Mil89}) $X = \res a (a. X | \outC a)$
where $X$ is guarded but not sequential. Any process that does not use
$a$ is a solution\hl{, e.g. $\nil$ and $b.\nil$.}

\hlD{For more details of above three theorems,} see Section~\ref{ss:part2}
for the univariate case and
Section~\ref{sec:multivariate} for the multivariate case.
%(Theorem~\ref{t:Mil89s1} only).

### 2019-09-14 ic-journal/expansion.tex
text

\hlD{\emph{Strong} bisimulation and strong bisimilarity} ($\sim$)
can be obtained by replacing the weak transition $Q\Arcap\mu Q'$ in the above
definition with normal transitions $Q \arr \mu Q'$ (similar for the other case).
Weak bisimilarity is not preserved by the sum operator (except for
guarded sums), \hlD{i.e.\  $P_1 \wb Q_1$ and $P_2 \wb Q_2$  does not imply
 $P_1 + P_2 \wb Q_1 + Q_2$.}
For this, Milner introduced the concept of \emph{observational congruence}, also called \emph{rooted
  bisimilarity}~\cite{Gorrieri:2015jt,Sangiorgi:2011ut}:
\begin{definition}%[rooted bisimilarity]
\label{d:rootedBisimilarity}
Two processes $P$ and $Q$ are \emph{rooted bisimilar}, written as $P
\rapprox Q$, \hl{if} % for all $\mu$:
%  for all $\mu\in \mathscr{L}\cup\{\tau\}$
\begin{enumerate}
 \item  $P \arr\mu P'$ implies that there is $Q'$ such that $Q
   \Arr\mu Q'$ and $P' \wb Q'$;
 \item  $Q \arr\mu Q'$ implies that there is $P'$ such that $P
   \Arr\mu P'$ and $P' \wb Q'$\enspace.
\end{enumerate}
\end{definition}

@@

\hlD{One basic property of rooted bisimiarity is congruence,
indeed
 the
coarsest congruence contained in bisimilarity
(c.f., Section~\ref{s:coarsest} for more details): }
\begin{theorem}
\label{t:rapproxCongruence}
$\rapprox$ is a congruence in CCS, and it is the
coarsest congruence contained in $\wb$.
\end{theorem}

>>>>

\hlD{\emph{Strong} bisimulation and strong bisimilarity} ($\sim$)
can be obtained by replacing the weak transition $Q\Arcap\mu Q'$ in the above
definition with normal transitions $Q \arr \mu Q'$ (similar for the other case).
Weak bisimilarity is not preserved by the sum operator (except for
guarded sums), i.e.\  $P_1 \wb Q_1$ and $P_2 \wb Q_2$  \hlD{does not imply}
 $P_1 + P_2 \wb Q_1 + Q_2$.
For this, Milner introduced the concept of \emph{observational congruence}, also called \emph{rooted
  bisimilarity}~\cite{Gorrieri:2015jt,Sangiorgi:2011ut}:
\begin{definition}%[rooted bisimilarity]
\label{d:rootedBisimilarity}
Two processes $P$ and $Q$ are \emph{rooted bisimilar}, written as $P
\rapprox Q$, \hl{if} % for all $\mu$:
%  for all $\mu\in \mathscr{L}\cup\{\tau\}$
\begin{enumerate}
 \item  $P \arr\mu P'$ implies that there is $Q'$ such that $Q
   \Arr\mu Q'$ and $P' \wb Q'$;
 \item  $Q \arr\mu Q'$ implies that there is $P'$ such that $P
   \Arr\mu P'$ and $P' \wb Q'$\enspace.
\end{enumerate}
\end{definition}

@@

\hlD{One basic property of rooted bisimiarity is congruence,
indeed
 the
coarsest congruence contained in bisimilarity}
(c.f., Section~\ref{s:coarsest} \hlD{for more details): }
\begin{theorem}
\label{t:rapproxCongruence}
$\rapprox$ is a congruence in CCS, and it is the
coarsest congruence contained in $\wb$.
\end{theorem}

### 2019-09-14 ic-journal/formal.tex
Moved context.htex after bisim.htex and before part2.htex (unique-solution theorems)
text

We highlight here a formalisation of CCS in the HOL theorem prover
(HOL4)~\cite{Melham:1993vl,slind2008brief}, \hl{with a focus towards the
theory (and formal proofs) of the unique solution of
equations/contractions theorems} mentioned in Section~\ref{s:eq} and
\ref{s:mcontr}.
%  The main purpose is to convince the readers that, there's no flaw
%  in the informal proofs. 
\hl{The univariate versions of these theorems, apart from minor fixes and
extensions in this paper, was already described in}
\cite{Tian:2017wrba}. The multivariate proofs and the related results
are new, and will be described in Section~\ref{sec:multivariate}.
All proof scripts are available as part of HOL's official
examples\footnote{\url{https://github.com/HOL-Theorem-Prover/HOL/tree/master/examples/CCS}}.
The work so far consists of about 24,000 lines (1MB) of code in total,
in which about 5,000 lines were derived from the early work of Monica
Nesi~\cite{Nesi:1992ve} on HOL88, with major modifications.

Higher Order Logic (HOL)~\cite{hollogic} traces
its roots back to LCF
\cite{gordon1979edinburgh,milner1972logic} by Robin Milner and others
since \hl{1972. It} is a variant of
Church’s Simple Theory of Types (STT)~\cite{church1940formulation},
plus a higher order version of Hilbert's choice operator $\varepsilon$,
Axiom of Infinity, and Rank-1 (prenex) polymorphism.
HOL4 has implemented the original HOL, 
while some other theorem provers in HOL family (e.g. Isabelle/HOL) have
certain extensions.
%  (they made the formal language more powerful,
% but they also bring the possibilities that the entire logic becomes
% inconsistent). 
Indeed the HOL has \hl{considerably} simpler logical
foundations than most other theorem provers. %, e.g. Coq. 
\hl{As a consequence, theories and proofs verified in HOL are relatively more
convincible to non-experts who are not familar with more advanced
dependent type theories.}

HOL4 is implemented in Standard ML, the same programming language who
plays three different roles:
\begin{enumerate}
\item the underlying implementation language for the core HOL engine;
\item \hl{the language in which proof tactics are implemented;}
\item \hl{the interface (outer) language of the HOL proof scripts and
  interactive shell.}
\end{enumerate}
Moreover, using the same language HOL4 users can write complex automatic
verification tools by calling HOL's theorem proving
facilities. The formal proofs of theorems in CCS theory
are mostly done by an \emph{interactive process} closely following
their informal proofs, with minimal automatic proof searching.

@@

Finally, the whole HOL \emph{standard} theory is based on the
following four axioms, from which the almost\footnote{HOL is strictly
  weaker than ZFC (the Zermelo-Frankel set theory with the
Axiom of Choice), thus not all theorems valid in ZFC can be formalised
in HOL. (c.f.~\cite{hollogic} for more details.)} entire mathematics
can be formalised:
\begin{equation*}
\begin{array}{@{}l@{\qquad}l}
\mbox{\texttt{BOOL\_CASES\_AX}} &\vdash \uquant{b} (b = \T ) \vee (b = \F )\\
\mbox{\texttt{ETA\_AX}} &
\vdash \uquant{f_{\alpha\fun\beta}}(\lquant{x}f\ x) = f\\
\mbox{\texttt{SELECT\_AX}} &
\vdash \uquant{P_{\alpha\fun\ty{bool}}\ x} P\ x \imp P({\hilbert}\ P)\\
\mbox{\texttt{INFINITY\_AX}}&
\vdash \equant{f_{\ind\fun \ind}} \OneOne \ f \conj \neg(\Onto \ f)\\
\end{array}
\end{equation*}

>>>>

We highlight here a formalisation of CCS in the HOL theorem prover
(HOL4)~\cite{Melham:1993vl,slind2008brief}, \hl{with a focus towards the
theory (and formal proofs) of the unique solution of
equations/contractions theorems} mentioned in Section~\ref{s:eq} and
\ref{s:mcontr}.
%  The main purpose is to convince the readers that, there's no flaw
%  in the informal proofs. 
% \hl{The univariate versions of these theorems, apart from minor fixes and
% extensions in this paper, was already described in}
% \cite{Tian:2017wrba}. The multivariate proofs and the related results
% are new, and will be described in Section~\ref{sec:multivariate}.
All proof scripts are available as part of HOL's official
examples\footnote{\url{https://github.com/HOL-Theorem-Prover/HOL/tree/master/examples/CCS}}.
The work so far consists of about 24,000 lines (1MB) of code in total,
in which about 5,000 lines were derived from the early work of Monica
Nesi~\cite{Nesi:1992ve} on HOL88, with major modifications.

Higher Order Logic (HOL)~\cite{hollogic} traces
its roots back to LCF
\cite{gordon1979edinburgh,milner1972logic} by Robin Milner and others
since \hl{1972. It} is a variant of
Church’s Simple Theory of Types (STT)~\cite{church1940formulation},
plus a higher order version of Hilbert's choice operator $\varepsilon$,
Axiom of Infinity, and Rank-1 (prenex) polymorphism.
HOL4 has implemented the original HOL, 
while some other theorem provers in HOL family (e.g. Isabelle/HOL) have
certain extensions.
%  (they made the formal language more powerful,
% but they also bring the possibilities that the entire logic becomes
% inconsistent). 
Indeed the HOL has \hl{considerably} simpler logical
foundations than most other theorem provers. %, e.g. Coq. 
\hl{As a consequence, theories and proofs verified in HOL are easier to understand
 to people who are not familar with more advanced
dependent type theories.}

HOL4 is implemented in Standard ML, the same programming language who
plays three different roles:
\begin{enumerate}
\item the underlying implementation language for the core HOL engine;
\item \hl{the language in which proof tactics are implemented;}
\item \hl{the interface  language of the HOL proof scripts and
  interactive shell.}
\end{enumerate}
Moreover, using the same language HOL4 users can write complex automatic
verification tools by calling HOL's theorem proving
facilities. The formal proofs of theorems in CCS theory
are mostly done by an \emph{interactive process} closely following
their informal proofs, with minimal automatic proof searching.

@@

Finally, the whole HOL \emph{standard} theory is based on the
following four axioms, from which the almost\footnote{HOL is strictly
  weaker than ZFC (the Zermelo-Frankel set theory with the
Axiom of Choice), thus not all theorems valid in ZFC can be formalised
in HOL. (c.f.~\cite{hollogic} for more details.)} entire mathematics
can be formalised:
\begin{equation*}
\begin{array}{@{}l@{\qquad}l}
\mbox{\texttt{BOOL\_CASES\_AX}} &\vdash \uquant{b} (b = \T )\ \vee \ (b = \F )\\
\mbox{\texttt{ETA\_AX}} &
\vdash \uquant{f_{\alpha\fun\beta}}(\lquant{x}f\ x) = f\\
\mbox{\texttt{SELECT\_AX}} &
\vdash \uquant{P_{\alpha\fun\ty{bool}}\ x} P\ x \imp P({\hilbert}\ P)\\
\mbox{\texttt{INFINITY\_AX}}&
\vdash \equant{f_{\ind\fun \ind}} \OneOne \ f \conj \neg(\Onto \ f)\\
\end{array}
\end{equation*}

### 2019-09-14 ic-journal/intro.tex
text

From the \hl{point of} view of the CCS theory, this formalisation has offered us the possibility of
further refining the theory of unique solutions of contractions.
In particular, \hlD{the existing result~\cite{sangiorgi2017equations} has
 limitations} on the body of the contractions due to the
substitutivity problems of weak bisimilarity and other behavioural relations with respect
to the sum operator.
\hl{In this paper, the contraction-based proof technique is further
refined by moving from the contraction preorder} to
\emph{rooted contraction}, that is, the coarsest precongruence contained in the contraction
preorder. The resulting unique-solution theorem is now valid for
\emph{rooted bisimilarity} (hence also for bisimilarity itself), and places no 
\hlD{constraints on the use the sum operator}.

@@

\hlD{We first consider the case of a single equation (or contraction),
that we call the univariate case, then consider the multivariate
  case, where  more than one equation
  (or contraction) is possible.
The univariate  theorems use
$\lambda$-functions to present  CCS equations,
while the multivariate versions require a
more careful and delicate treatment of 
expressions with several 
variables and substitutions acting on them.
In contrast to most literature, we have adopted the same
type for both CCS equations and processes: those undefined constants
in CCS terms are treated as  (free) equation variables, and a CCS
\emph{process} is a CCS term without equation variables. 
This allows us a smoother move from the univariate case to the multivariate
one. See Section~\ref{sec:multivariate}
for  deails.}

@@

Section~\ref{ss:ccs} presents basic background materials on CCS,
including its syntax, operational semantics, bisimilarity and rooted
bisimilarity.
Section~\ref{s:eq} discusses equations and contractions.
Section~\ref{ss:new} presents rooted contraction and the related
unique-solution result for rooted bisimilarity.
Section~\ref{s:for} highlights our formalisation in HOL4 \hlDfor the
univariate case.}
\hlD{Section~\ref{sec:multivariate} describes the move to the
  multivariate case}. 
Finally,  Section~\ref{s:rel} and \ref{s:concl} discuss related work,
conclusions, and a few directions for future work.

>>>>

From the \hl{point of} view of the CCS theory, this formalisation has offered us the possibility of
further refining the theory of unique solutions of contractions.
In particular, \hlD{the existing result}~\cite{sangiorgi2017equations} \hlD{has  limitations} on the body of the contractions due to the
substitutivity problems of weak bisimilarity and other behavioural relations with respect
to the sum operator.
\hl{In this paper, the contraction-based proof technique is further
refined by moving from the contraction preorder} to
\emph{rooted contraction}, that is, the coarsest precongruence contained in the contraction
preorder. The resulting unique-solution theorem is now valid for
\emph{rooted bisimilarity} (hence also for bisimilarity itself), and places no 
\hlD{constraints on the use the sum operator}.

@@

\hlD{We first consider the case of a single equation (or contraction),
that we call the univariate case, then consider the multivariate
  case, where  more than one equation
  (or contraction) is possible.
The univariate  theorems use
$\lambda$-functions to present  CCS equations,
while the multivariate versions require a
more careful and delicate treatment of 
expressions with several 
variables and substitutions acting on them.
In contrast to most literature, we have adopted the same
type for both CCS equations and processes: those undefined constants
in CCS terms are treated as  (free) equation variables, and a CCS
\emph{process} is a CCS term without equation variables. 
This allows us a smoother move from the univariate case to the multivariate
one. See} Section~\ref{sec:multivariate} for  details.

@@

Section~\ref{ss:ccs} presents basic background materials on CCS,
including its syntax, operational semantics, bisimilarity and rooted
bisimilarity.
Section~\ref{s:eq} discusses equations and contractions.
Section~\ref{ss:new} presents rooted contraction and the related
unique-solution result for rooted bisimilarity.
Section~\ref{s:for} highlights our formalisation in HOL4 \hlD{for the
univariate case.}
Section~\ref{sec:multivariate} \hlD{describes the move to the
  multivariate case}. 
Finally,  Section~\ref{s:rel} and \ref{s:concl} discuss related work,
conclusions, and a few directions for future work.
